---
title: SLICED s01e02
author: Jo Dudding
output: 
  html_document:
    theme: united
    highlight: kate
    toc: true
    toc_float: 
      collapsed: true
    includes:
      in_header: "header.html"
      after_body: "footer.html"

---

```{r setup, include = FALSE, echo = FALSE}
#-------------------------------------------------------------------------------
#' set knitr options
#-------------------------------------------------------------------------------

knitr::opts_chunk$set(
  fig.width = 24 / 2.54,
  fig.height = 13.4 / 2.54,
  fig.retina = 4,
  comment = '',
  warning = FALSE
)
```

```{css, echo = FALSE}
h1, h2, h3, h4, h5, h6, legend {
  color: #302df0;
}

h1.title {
    font-size: 38px;
    background-color: #e946a5;
    color: #ffffff;
    margin-bottom: 20px;
    padding: 9.5px;
    padding-top: 20px;
    padding-bottom: 20px;
    border-radius: 10px;
}

pre {
  background-color: #cccccc;
}

a {
  color: #129875;
}

.hljs-comment {
    color: #68e199;
    font-style: 68e199;
}

.list-group-item.active, .list-group-item.active:hover, .list-group-item.active:focus {
    z-index: 2;
    color: #ffffff;
    background-color: #68e199;
    border-color: #68e199;
}
```

Predict whether an aircraft strike with wildlife causes damage

[Kaggle link](https://www.kaggle.com/c/sliced-s01e02-xunyc5)

[Twitch link](https://www.twitch.tv/nickwan_datasci)

This time I'm trying an XGBoost model.

# Load libraries

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(tidymodels)
library(scales)
library(skimr)
library(kableExtra)
library(patchwork)
library(gt)
library(lubridate)
library(glue)
library(themis)
```

# My settings

```{r}

# colours
#pal_jo <- c(viridisLite::magma(8)[2:8], "#4C4C53", "#9B9BA8")

pal_sliced <- c(
  '#e946a5', # pink
  '#68e199', # green
  '#302df0', # blue
  '#6a45b0', # purple
  '#129875', # darker green
  '#4f94bf', # lighter blue
  '#b545ab', # darker pink
  '#000000', # black,
  '#cccccc'  #grey
)

title_colour <- pal_sliced[1]
table_colour <- pal_sliced[4]
bar_colour <- pal_sliced[2]
grid_colour <- pal_sliced[9]

theme_set(theme_bw() %+replace%
    theme(
      # align title and caption to the plot not the panel
      plot.title.position = 'plot',
      plot.caption.position = 'plot',
      # change the title and caption to markdown and move them futher from the plot
      plot.title = element_text(
        size = rel(1.3),
        hjust = 0, 
        margin = margin(c(0, 0, 10, 0)),
        colour = title_colour
      ),
      plot.subtitle = element_text(
        size = rel(1.15),
        hjust = 0, 
        margin = margin(c(0, 0, 15, 0))
      ),
      plot.caption = element_text(
        hjust = 1, 
        margin = margin(c(10, 0, 0, 0))
      ),
      # move axis titles to the left/top and change them to markdown
      axis.title = element_text(hjust = 1),
      # allow the axis values to the markdown as well
      axis.text = element_text(),
      # remove the panel border
      panel.border = element_blank(),
      # put in the axis lines with a slightly thicker line than the gridlines
      axis.line = element_line(colour = grid_colour, size = rel(1.5)),
      # make the tickmarks the same colour
      axis.ticks = element_line(colour = grid_colour),
      # facet strip text left aligned with extra space above
      strip.text = element_text(
        hjust = 0, margin = margin(c(10, 0, 0, 0)), colour = title_colour
      ),
      # clear colour and fill for strip
      strip.background = element_rect(colour = NA, fill = NA),
      # dotted gridlines
      panel.grid = element_line(linetype = 'dotted'),
      # ability to use a different colour for the gridlines
      panel.grid.major.x = element_line(colour = grid_colour),
      panel.grid.major.y = element_line(colour = grid_colour),
      panel.grid.minor.x = element_blank(),
      panel.grid.minor.y = element_blank(),
    )
)

scale_y_pct <- function(
  accuracy = 1L, 
  breaks = pretty_breaks(),
  expand = expansion(mult = c(0, .05)),
  ...
) {
  scale_y_continuous(
    labels = scales::percent_format(accuracy = accuracy, big.mark = ","),
    breaks = breaks,
    expand = expand,
    ...
  )
}

scale_y_comma <- function(
  accuracy = 1L, 
  breaks = pretty_breaks(),
  expand = expansion(mult = c(0, .05)),
  ...
) {
  scale_y_continuous(
    labels = scales::comma_format(accuracy = accuracy),
    breaks = breaks,
    expand = expand,
    ...
  )
}

scale_fill_jo <- function(...) {
  scale_fill_manual(values = pal_sliced, ...)
}

scale_fill_discrete <- scale_fill_jo

update_geom_defaults("bar", list(fill = bar_colour))
update_geom_defaults("col", list(fill = bar_colour))
update_geom_defaults("point", list(colour = bar_colour))
update_geom_defaults("line", list(colour = bar_colour))
```


# Read files

```{r}
to_build <- read_csv("s01e02/train.csv", guess_max = 200000) 
to_score <- read_csv("s01e02/test.csv", guess_max = 200000) 
```

# Examine data

```{r}
skim(to_build)
```

<br>
Initial thoughts

- Lots of missing values!
- Some of the character variables have lots of categories
- Missing value for the target variable

```{r}
explore_build <- to_build %>% 
  filter(! is.na(damaged)) %>% 
  mutate(
    damaged_factor = factor(if_else(damaged == 1, 'Damaged', 'Not Damaged')),
    date = ymd(paste(incident_year, incident_month, incident_day, 
      sep = '-')),
    dow = wday(date, label = TRUE, abbr = TRUE),
    log_height= log(height),
    log_speed= log(speed),
    log_distance = log(distance),
    aircraft_mass = factor(aircraft_mass)
  )

```

## Look at sample

```{r}
set.seed(617)

explore_build %>% 
  select(damaged, damaged_factor, everything()) %>% 
  slice_sample(n = 5) %>% 
  gather(-id, key = 'key', value = 'value') %>% 
  spread(key = id, value = value) %>% 
  arrange(-str_detect(key, 'damag')) %>% 
  gt()

```

## Target - damaged

- unbalanced
- was a missing value

```{r}
summary(explore_build$damaged)

ggplot(explore_build, aes(damaged_factor, fill = damaged_factor)) +
  geom_bar() +
  scale_y_comma() +
  labs(title = 'Damaged', x = NULL, y = NULL, fill = 'Damaged') +
  guides(fill = 'none')
```


## Function for exploration - numeric


```{r}

explore_num <- function(x, ...){

a <- explore_build %>% 
  mutate(is_it_na = if_else(is.na({{x}}), 'NA', 'Present')) %>% 
ggplot(aes(is_it_na, fill = damaged_factor)) +
  geom_bar() +
  scale_y_comma() +
  labs(x = NULL, y = NULL, fill = 'Damaged') +
  guides(fill = 'none')

b <- explore_build %>% 
  mutate(is_it_na = if_else(is.na({{x}}), 'NA', 'Present')) %>% 
ggplot(aes(is_it_na, fill = damaged_factor)) +
  geom_bar(position = 'fill') +
  scale_y_pct() +
  labs(x = NULL, y = NULL, fill = 'Damaged') +
  guides(fill = 'none')

c <- ggplot(explore_build, aes({{x}}, fill = damaged_factor)) +
  geom_histogram( ...) +
  scale_y_comma() +
  labs(x = NULL, y = NULL, fill = 'Damaged') +
  guides(fill = 'none')

d <- ggplot(explore_build, aes({{x}}, fill = damaged_factor)) +
  geom_histogram(position = 'fill', ...) +
  scale_y_pct() +
  labs(x = NULL, y = NULL, fill = 'Damaged') +
  guides(fill = 'none')

(a + b) / (c + d) 

}

```

## Function for exploration - category


```{r}

explore_char <- function(x, ...){

c <- explore_build %>% 
  mutate(
    tmp = fct_infreq(fct_lump_n({{x}}, 20))
  ) %>% 
ggplot(aes(tmp, fill = damaged_factor)) +
  geom_bar() +
  scale_y_comma() +
  scale_x_discrete(label = label_wrap(15)) +
  labs(x = NULL, y = NULL, fill = 'Damaged') +
  coord_flip() +
  guides(fill = 'none')

d <- explore_build %>% 
  mutate(
    tmp = fct_infreq(fct_lump_n({{x}}, 20))
  ) %>% 
ggplot(aes(tmp, fill = damaged_factor)) +
  geom_bar(position = 'fill') +
  scale_y_pct() +
  scale_x_discrete(label = label_wrap(15)) +
  labs(x = NULL, y = NULL, fill = 'Damaged') +
  coord_flip() +
  guides(fill = 'none')

(c + d) 
}

```

## incident_year

Year the incident occurred (1990-2015)

- No missing data
- Rate decreasing over time


```{r}
explore_num(incident_year, binwidth = 1)
```  

## incident_month

Month the incident occurred (1-12)

- Change to factor
- seasonal impact


```{r}
explore_num(incident_month, binwidth = 1)
``` 

## incident_day

Day the incident occurred (1-31)

- Omit raw
- Could combine with month and year to get date and use `step_date`?
- Public holidays could be an issue `step_holiday`
- Day of week might be more useful - Not useful


```{r}
explore_char(dow)
``` 

## operator_id / operator

International Civil Aviation Organization (ICAO) code for aircraft operators / Name of the Airline

- Omit operator_id
- some operators have higher rates

```{r}
explore_char(operator)
``` 

## aircraft

Model of aircraft involved in the strike

For example: B-777 means a Boeing 777 was struck in the incident

```{r}
explore_char(aircraft)
``` 

## aircraft_type

- 'A': "multi-engined aeroplanes powered by turbo-propeller engines with an MOPSC of more than nine or a maximum take-off mass exceeding 5 700 kg, and all multi-engined turbo-jet powered aeroplanes"
- 'B': "aeroplanes powered by propeller engines with an MOPSC of nine or less and a maximum take-off mass of 5 700 kg or less"

```{r}
explore_char(aircraft_type)
``` 

## aircraft_make

International Civil Aviation Organization (ICAO) code for Aircraft Make

```{r}
explore_char(aircraft_make)
``` 

## aircraft_model

International Civil Aviation Organization (ICAO) code for Aircraft Model

- should be factor

```{r}
explore_num(aircraft_model, binwidth = 10)
``` 

## aircraft_mass

- 1 = 2,250 kg or less
- 2 = 2,251-5,700 kg
- 3 = 5,701-27,000 kg
- 4 = 27,001-272,000 kg
- 5 = above 272,000 kg

- convert to factor

```{r}
explore_char(aircraft_mass)
``` 

## engine_make

Engine Make Code

- should be a factor

```{r}
explore_num(engine_make, binwidth = 1)
``` 

## engine_model

Engine Model Code

```{r}
explore_char(engine_model)
``` 

## engines

Number of engines powering the aircraft

```{r}
explore_num(engines, binwidth = 1)
``` 

## engine_type

The way the engine is powered

- 'A' - Reciprocating
- 'B' - Turbojet
- 'C' - Turboprop
- 'D' - Turbofan
- 'E' - None (glider)
- 'F' - Turboshaft (helicopter)
- Special Note: Some aircrafts can have two engine types
  - Example: 'B/D'

```{r}
explore_char(engine_type)
``` 

## engine1_position

Where engine 1 is mounted on aircraft

- should be  factor

```{r}
explore_num(engine1_position)
``` 

## engine2_position

Where engine 2 is mounted on aircraft

- should be a factor

```{r}
explore_num(engine2_position)
``` 

## engine3_position

Where engine 3 is mounted on aircraft

- omit

## engine4_position

Where engine 4 is mounted on aircraft

- omit

## airport_id / airport

Location of the strike given as the International Civil Aviation Organization (ICAO) airport identifier / Name of the airport where the strike occurred

```{r}
explore_char(airport)
``` 

## state

Abbreviated state from US, Canada

```{r}
explore_char(state)
``` 

## faa_region
FAA region where the airport is located

- The US has 9 regions (AAL, ACE, AEA, AGL, ANE, ANM, ASO, ASW, AWP)
- Canada has 5 regions (ATL, AWP, ONT, P&N, PAC, QUE)
- All other regions of the world are considered foreign regions (FGN)

```{r}
explore_char(faa_region)
``` 

## flight_phase

Phase of flight during which the strike occured

- APPROACH
- ARRIVAL
- CLIMB
- DEPARTURE
- DESCENT
- EN ROUTE
- LANDING
- LANDING ROLL
- LOCAL
- PARKED
- TAKEOFF RUN
- TAXI

```{r}
explore_char(flight_phase)
```


## visibility

Time of day when the strike occurred

- DAWN
- DAY
- DUSK
- NIGHT
- UNKNOWN

```{r}
explore_char(visibility)
```

## precipitation

- FOG
 -RAIN
- SNOW
- NONE
- Special Note: Can have multiple values
  - Example: FOG, RAIN
  
- Is it worth splitting?  

```{r}
explore_char(precipitation)
```

## height

Feet About Ground Level where the strike occurred (0-24,000)

```{r}
explore_num(height, bins = 50)
``` 

Try log

```{r}
explore_num(log_height, bins = 50)
``` 

## speed

Velocity in Knots (0-2500)

- trim high values

```{r}
explore_num(speed, bins = 50)
``` 

Try log

```{r}
explore_num(log_speed, bins = 50)
``` 

## distance

How many miles from the airport (0-100)

```{r}
explore_num(distance, binwidth = 5)
``` 

Try log

```{r}
explore_num(log_distance, bins = 50)
``` 

## species_id / species_name

International Civil Aviation Organization (ICAO) code for type of wildlife / Common name for type of wildlife struck

- just use name

```{r}
explore_char(species_name)
```

## species_quantity

Amount of wildlife struck

- 1
- 2-10
- 11-100
- Over 100


```{r}
explore_char(species_quantity)
```

## flight_impact

How the strike impacted the flight

- ABORTED
- TAKEOFF
- ENGINE SHUTDOWN
- NONE
- OTHER
- PRECAUTIONARY LANDING

```{r}
explore_char(flight_impact)
```

# Split in to train and test

```{r}
set.seed(728)

to_build <- to_build %>% 
  mutate(damaged = factor(damaged, levels = c(1, 0)))

split <- initial_split(to_build, strata = damaged)
train <- training(split)
test <- testing(split)
```

# Create folds

```{r}
set.seed(530)
train_folds <- vfold_cv(train, 5)
```


# Create recipe

```{r}
rec <- recipe(damaged ~ ., data = train) %>%
  #recipe(damaged ~ speed + height + distance + incident_year + incident_month,
  #  data = train) %>% 
  # id role
  update_role(id, new_role = "ID") %>%
  # damaged not missing
  step_filter(! is.na(damaged)) %>% 
  # create variables
  step_mutate(
    speed = if_else(speed > 500, 500, speed)
  ) %>% 
  # numbers to characters
  step_mutate(
    incident_month = as.character(incident_month), 
    aircraft_mass = as.character(aircraft_mass), 
    aircraft_model = as.character(aircraft_model), 
    engine_make = as.character(engine_make), 
    engine1_position = as.character(engine1_position), 
    engine2_position = as.character(engine2_position),
    engines = as.character(engines)
  ) %>% 
  # remove variables
  step_rm(operator_id, incident_day, species_id, airport_id, 
    engine3_position, engine4_position) %>%
  # flag for missing values
  step_indicate_na(height, speed, distance) %>% 
  # imputation of missing values
  step_impute_mean(height, speed, distance) %>% 
  # character to factors
  step_string2factor(all_nominal()) %>% 
  # explicit na
  step_unknown(all_nominal_predictors()) %>% 
  # handle new classes
  step_novel(all_nominal(), -all_outcomes()) %>% 
  # lump together small values
  step_other(all_nominal_predictors()) %>% 
  # all numeric
  step_dummy(all_nominal_predictors()) %>% 
  # missing values for factors
  step_impute_lower(all_numeric_predictors()) %>% 
  # remove if zero variance
  step_zv(all_predictors()) %>% 
  # rebalance
  step_downsample(damaged)
```

# Check recipe outputs

```{r}
prepped <- prep(rec)
baked <- bake(prepped, new_data = NULL) 

skim(baked)
```
<br>


# Specify model

```{r}
xgboost_spec <- boost_tree(
    trees = 2000, 
    min_n = 23, 
    tree_depth = 15, 
    learn_rate = 0.00591673615298013, 
    loss_reduction = 3.41525742335076e-07, 
    sample_size = 0.648717504832894
  ) %>% 
  set_mode("classification") %>% 
  set_engine("xgboost")
```

# Specify workflow

```{r}
xgboost_workflow <- workflow() %>% 
  add_recipe(rec) %>% 
  add_model(xgboost_spec) 

xgboost_workflow
```
# Tune

Tuning happened but won't be included because it took so long.  

```{r, eval = FALSE}
set.seed(34379)
xgboost_tune <- tune_grid(
  xgboost_workflow, 
  resamples = train_folds, 
  grid = 9
 )
```

## Tuning outcome

The results were:

- trees = 2000
- min_n = 23
- tree_depth = 15
- learn_rate = 0.00591673615298013
- loss_reduction = 3.41525742335076e-07
- sample_size = 0.648717504832894


```{r, eval = FALSE}
show_best(xgboost_tune, metric = "roc_auc") %>%
  gt()

autoplot(xgboost_tune) +
  theme_bw()
```

# Finalise model

```{r}
final_wkflow <- xgboost_workflow #%>%
  #finalize_workflow(select_best(xgboost_tune, metric = "roc_auc"))

final_wkflow
```

# Fit model

```{r}
xgboost_fit <- last_fit(final_wkflow, split)
```

# Check performance

```{r}
collect_metrics(xgboost_fit) %>%
  gt()
```

# Get predictions

```{r}
xgboost_preds <- collect_predictions(xgboost_fit)
```

# ROC Curve

```{r}
xgboost_preds %>%
  roc_curve(damaged, .pred_1) %>% 
  autoplot()
```

# Apply to new data

```{r}
have_scored <- cbind(to_score, predict(xgboost_fit$.workflow[[1]], to_score)) %>%
  select(id, damaged = .pred)
```

# Output submission

```{r}
write_csv(have_scored, 'sliced/s01e02/submission.csv')
```

# Outcome



# Possible improvements

- 
