---
title: SLICED s01e01
author: Jo Dudding
output: 
  html_document:
    theme: united
    highlight: kate
    toc: true
    toc_float: 
      collapsed: true
    includes:
      in_header: "../sliced/header.html"
      after_body: "../sliced/footer.html"

---

```{r setup, include = FALSE, echo = FALSE}
#-------------------------------------------------------------------------------
#' set knitr options
#-------------------------------------------------------------------------------

knitr::opts_knit$set(
  root.dir = ".."
)

knitr::opts_chunk$set(
  fig.width = 19 / 2.54,
  fig.height = 13.4 / 2.54,
  fig.retina = 4,
  comment = '',
  warning = FALSE
)
```

```{css, echo = FALSE}
h1, h2, h3, h4, h5, h6, legend {
  color: #302df0;
}

h1.title {
    font-size: 38px;
    background-color: #e946a5;
    color: #ffffff;
    margin-bottom: 20px;
    padding: 9.5px;
    padding-top: 20px;
    padding-bottom: 20px;
    border-radius: 10px;
}

pre {
  background-color: #cccccc;
}

a {
  color: #129875;
}

.hljs-comment {
    color: #68e199;
    font-style: 68e199;
}

.list-group-item.active, .list-group-item.active:hover, .list-group-item.active:focus {
    z-index: 2;
    color: #ffffff;
    background-color: #68e199;
    border-color: #68e199;
}
```

Predict the popularity of board games on BoardGameGeek.com

[Kaggle link](https://www.kaggle.com/c/sliced-s01e01/overview/description)

[Twitch link](https://www.twitch.tv/nickwan_datasci)

# Load libraries

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(tidymodels)
library(scales)
library(skimr)
library(kableExtra)

# colours
#pal_jo <- c(viridisLite::magma(8)[2:8], "#4C4C53", "#9B9BA8")

pal_sliced <- c(
  '#e946a5', # pink
  '#68e199', # green
  '#302df0', # blue
  '#6a45b0', # purple
  '#129875', # darker green
  '#4f94bf', # lighter blue
  '#b545ab', # darker pink
  '#000000', # black,
  '#cccccc'  #grey
)

title_colour <- pal_sliced[1]
table_colour <- pal_sliced[4]
bar_colour <- pal_sliced[2]
grid_colour <- pal_sliced[9]

theme_set(theme_bw() %+replace%
    theme(
      # align title and caption to the plot not the panel
      plot.title.position = 'plot',
      plot.caption.position = 'plot',
      # change the title and caption to markdown and move them futher from the plot
      plot.title = element_text(
        size = rel(1.3),
        hjust = 0, 
        margin = margin(c(0, 0, 10, 0)),
        colour = title_colour
      ),
      plot.subtitle = element_text(
        size = rel(1.15),
        hjust = 0, 
        margin = margin(c(0, 0, 15, 0))
      ),
      plot.caption = element_text(
        hjust = 1, 
        margin = margin(c(10, 0, 0, 0))
      ),
      # move axis titles to the left/top and change them to markdown
      axis.title = element_text(hjust = 1),
      # allow the axis values to the markdown as well
      axis.text = element_text(),
      # remove the panel border
      panel.border = element_blank(),
      # put in the axis lines with a slightly thicker line than the gridlines
      axis.line = element_line(colour = grid_colour, size = rel(1.5)),
      # make the tickmarks the same colour
      axis.ticks = element_line(colour = grid_colour),
      # facet strip text left aligned with extra space above
      strip.text = element_text(
        hjust = 0, margin = margin(c(10, 0, 0, 0)), colour = title_colour
      ),
      # clear colour and fill for strip
      strip.background = element_rect(colour = NA, fill = NA),
      # dotted gridlines
      panel.grid = element_line(linetype = 'dotted'),
      # ability to use a different colour for the gridlines
      panel.grid.major.x = element_line(colour = grid_colour),
      panel.grid.major.y = element_line(colour = grid_colour),
      panel.grid.minor.x = element_blank(),
      panel.grid.minor.y = element_blank(),
    )
)

scale_y_pct <- function(
  accuracy = 1L, 
  breaks = pretty_breaks(),
  expand = expansion(mult = c(0, .05)),
  ...
) {
  scale_y_continuous(
    labels = scales::percent_format(accuracy = accuracy, big.mark = ","),
    breaks = breaks,
    expand = expand,
    ...
  )
}

scale_y_comma <- function(
  accuracy = 1L, 
  breaks = pretty_breaks(),
  expand = expansion(mult = c(0, .05)),
  ...
) {
  scale_y_continuous(
    labels = scales::comma_format(accuracy = accuracy),
    breaks = breaks,
    expand = expand,
    ...
  )
}

scale_fill_jo <- function(...) {
  scale_fill_manual(values = pal_sliced, ...)
}

scale_fill_discrete <- scale_fill_jo

update_geom_defaults("bar", list(fill = bar_colour))
update_geom_defaults("col", list(fill = bar_colour))
update_geom_defaults("point", list(colour = bar_colour))
update_geom_defaults("line", list(colour = bar_colour))
```


# Read files

```{r}
to_build <- read_csv("sliced/s01e01/train.csv", guess_max = 200000) 
to_score <- read_csv("sliced/s01e01/test.csv", guess_max = 200000) %>%
  mutate(
    category11 = as.character(category11),
    category12 = as.character(category12)
  )
```

# Examine data

```{r}
skim(to_build)
```

## Target - geek rating

```{r}
summary(to_build$geek_rating)

ggplot(to_build, aes(geek_rating)) +
  geom_histogram(binwidth = 0.2) +
  scale_y_comma() +
  labs(title = 'Geek rating', x = NULL, y = NULL, fill = 'Rating > 7')
```

Score over 7 looks good

## Number of players 

- if zero reset to missing and impute
- if over 20 reset to 20

```{r}

summary(to_build$min_players)

ggplot(to_build, aes(min_players, fill = geek_rating >= 7)) +
  geom_histogram(binwidth = 1) +
  scale_y_comma() +
  labs(title = 'Min players', x = NULL, y = NULL, fill = 'Rating > 7')

ggplot(to_build, aes(min_players, fill = geek_rating >= 7)) +
  geom_histogram(binwidth = 1, position = 'fill') +
  scale_y_pct() +
  labs(title = 'Min players', x = NULL, y = NULL, fill = 'Rating > 7')

summary(to_build$max_players)

ggplot(to_build, aes(max_players, fill = geek_rating >= 7)) +
  geom_histogram(binwidth = 1) +
  xlim(0, 20) +
  scale_y_comma() +
  labs(title = 'Max players', x = NULL, y = NULL, fill = 'Rating > 7')

ggplot(to_build, aes(max_players, fill = geek_rating >= 7)) +
  geom_histogram(binwidth = 1, position = 'fill') +
  xlim(0, 20) +
  scale_y_pct() +
  labs(title = 'Max players', x = NULL, y = NULL, fill = 'Rating > 7')
```

## Year

- if zero then impute
- otherwise if < 1900 then set to 1900

```{r}
summary(to_build$year)

ggplot(to_build, aes(year, fill = geek_rating >= 7)) +
  geom_histogram(binwidth = 10) +
  scale_y_comma() +
  labs(title = 'Year', x = NULL, y = NULL, fill = 'Rating > 7')

to_build %>%
  filter(year < 0) %>%
  arrange(year) %>%
  select(names, year) %>%
  kable( 
    escape = FALSE,
    padding = 5
  ) %>% 
  kable_styling(
    bootstrap_options = c("striped", "hover"), 
    fixed_thead = TRUE, 
    position = "left"
  ) %>%
  row_spec(0, color = "white", background = table_colour)

to_build %>%
  filter(year >= 0 & year < 1900) %>%
  count(year) %>%
  kable( 
    escape = FALSE,
    padding = 5
  ) %>% 
  kable_styling(
    bootstrap_options = c("striped", "hover"), 
    fixed_thead = TRUE, 
    position = "left"
  ) %>%
  row_spec(0, color = "white", background = table_colour)


summary(to_build$year[to_build$year > 1900])

ggplot(to_build, aes(year, fill = geek_rating >= 7)) +
  geom_histogram(binwidth = 10)  +
  xlim(1900, 2020) +
  scale_y_comma() +
  labs(title = 'Year', x = NULL, y = NULL, fill = 'Rating > 7')

ggplot(to_build, aes(year, fill = geek_rating >= 7)) +
  geom_histogram(binwidth = 10, position = 'fill')  +
  xlim(1900, 2020) +
  scale_y_pct() +
  labs(title = 'Year', x = NULL, y = NULL, fill = 'Rating > 7')
```  

## Time

- if zero reset to missing and impute
- if > 240 then set to 240

```{r}

summary(to_build$avg_time)

ggplot(to_build, aes(avg_time, fill = geek_rating >= 7)) +
  geom_histogram(binwidth = 15) +
  xlim(0, 240) +
  scale_y_comma() +
  labs(title = 'Average Time', x = NULL, y = NULL, fill = 'Rating > 7')

ggplot(to_build, aes(avg_time, fill = geek_rating >= 7)) +
  geom_histogram(binwidth = 15, position = 'fill') +
  xlim(0, 240) +
  scale_y_pct() +
  labs(title = 'Average Time', x = NULL, y = NULL, fill = 'Rating > 7')  

summary(to_build$min_time)

ggplot(to_build, aes(min_time, fill = geek_rating >= 7)) +
  geom_histogram(binwidth = 15) +
  xlim(0, 240) +
  scale_y_comma() +
  labs(title = 'Min Time', x = NULL, y = NULL, fill = 'Rating > 7')

ggplot(to_build, aes(min_time, fill = geek_rating >= 7)) +
  geom_histogram(binwidth = 15, position = 'fill') +
  xlim(0, 240) +
  scale_y_pct() +
  labs(title = 'Min Time', x = NULL, y = NULL, fill = 'Rating > 7')

summary(to_build$max_time)

ggplot(to_build, aes(max_time, fill = geek_rating >= 7)) +
  geom_histogram(binwidth = 15) +
  xlim(0, 240) +
  scale_y_comma() +
  labs(title = 'Max Time', x = NULL, y = NULL, fill = 'Rating > 7')

ggplot(to_build, aes(max_time, fill = geek_rating >= 7)) +
  geom_histogram(binwidth = 15, position = 'fill') +
  xlim(0, 240) +
  scale_y_pct() +
  labs(title = 'Max Time', x = NULL, y = NULL, fill = 'Rating > 7')
```


## Number of votes

- log transform

```{r}
summary(to_build$num_votes)

ggplot(to_build, aes(num_votes, fill = geek_rating >= 7)) +
  geom_histogram(bins = 50) +
  scale_x_log10(label = comma_format()) +
  scale_y_comma() +
  labs(title = 'Number of votes', x = NULL, y = NULL, fill = 'Rating > 7')

ggplot(to_build, aes(num_votes, fill = geek_rating >= 7)) +
  geom_histogram(bins = 50, position = 'fill') +
  scale_x_log10(label = comma_format()) +
  scale_y_pct() +
  labs(title = 'Number of votes', x = NULL, y = NULL, fill = 'Rating > 7')
```

## Owned

- log transform

```{r}
summary(to_build$owned)

ggplot(to_build, aes(owned, fill = geek_rating >= 7)) +
  geom_histogram(bins = 50) +
  scale_x_log10(label = comma_format()) +
  scale_y_comma() +
  labs(title = 'Owned', x = NULL, y = NULL, fill = 'Rating > 7')

ggplot(to_build, aes(owned, fill = geek_rating >= 7)) +
  geom_histogram(bins = 50, position = 'fill') +
  scale_x_log10(label = comma_format()) +
  scale_y_pct() +
  labs(title = 'Owned', x = NULL, y = NULL, fill = 'Rating > 7')
```


## Age

- looks okay

```{r}
summary(to_build$age)

ggplot(to_build, aes(age, fill = geek_rating >= 7)) +
  geom_histogram(binwidth = 1) +
  xlim(0, 20) +
  scale_y_comma() +
  labs(title = 'Age', x = NULL, y = NULL, fill = 'Rating > 7')

ggplot(to_build, aes(age, fill = geek_rating >= 7)) +
  geom_histogram(binwidth = 1, position = 'fill')  +
  xlim(0, 20) +
  scale_y_pct() +
  labs(title = 'Age', x = NULL, y = NULL, fill = 'Rating > 7')
```

## Common categories

- flags where 5% +

```{r}
top_cat <- to_build %>%
  select(starts_with('cate')) %>%
  gather(key = 'key', value = 'value') %>%
  filter(! is.na(value)) %>%
  count(value, sort = TRUE) %>%
  #slice(1:20) %>%
  select(value, n) %>%
  mutate(pct = n /  nrow(to_build)) %>%
  filter(pct > 0.05)

ggplot(top_cat, aes(fct_reorder(value, n), pct)) +
  geom_col() +
  coord_flip() +
  scale_y_pct() +
  labs(title = 'Top categories', x = NULL, y = NULL, fill = 'Rating > 7')
```

## Common mechanic

- flags where 5% +

```{r}
top_mechanic <- to_build %>%
  select(mechanic) %>%
  mutate(mechanic = str_split(mechanic, ',')) %>%
  unnest(mechanic) %>%
  mutate(mechanic = str_trim(mechanic)) %>%
  count(mechanic, sort = TRUE) %>%
  #slice(1:20) %>%
  select(mechanic, n) %>%
  mutate(pct = n /  nrow(to_build)) %>%
  filter(pct > 0.05)

ggplot(top_mechanic, aes(fct_reorder(mechanic, n), pct)) +
  geom_col() +
  coord_flip() +
  scale_y_pct() +
  labs(title = 'Top mechanics', x = NULL, y = NULL, fill = 'Rating > 7')
```

## Designer

```{r}
top_designer <- to_build %>%
  select(designer)%>%
  mutate(designer = str_split(designer, ',')) %>%
  unnest(designer) %>%
  mutate(designer = str_trim(designer)) %>%
  count(designer, sort = TRUE) %>%
  filter(! designer %in% c('(Uncredited)', 'Jr.')) %>%
  slice(1:10) %>%
  select(designer, n) %>%
  mutate(pct = n /  nrow(to_build))

ggplot(top_designer, aes(fct_reorder(designer, n), pct)) +
  geom_col() +
  coord_flip()  +
  scale_y_pct(accuracy = 0.1) +
  labs(title = 'Top designers', x = NULL, y = NULL, fill = 'Rating > 7')
```

# Split in to train and test

```{r}
set.seed(539)

split <- initial_split(to_build, strata = geek_rating)
train <- training(split)
test <- testing(split)
```

# Create recipe

```{r}
rec <- recipe(geek_rating ~ ., data = train) %>%
  # id role for game_id
  update_role(game_id, new_role = "ID") %>%
  # flags for categories
  step_mutate(
    
    cat_card_game = if_any(starts_with("cate"), ~ .x == "Card Game"),
    cat_card_game =as.numeric( replace_na(cat_card_game, FALSE)),
    
    cat_wargame = if_any(starts_with("cate"), ~ .x == "Wargame"),
    cat_wargame = as.numeric(replace_na(cat_wargame, FALSE)),
    
    cat_fantasy = if_any(starts_with("cate"), ~ .x == "Fantasy"),
    cat_fantasy = as.numeric(replace_na(cat_fantasy, FALSE)),
    
    cat_economic = if_any(starts_with("cate"), ~ .x == "Economic"),
    cat_economic = as.numeric(replace_na(cat_economic, FALSE)),
    
    cat_fighting = if_any(starts_with("cate"), ~ .x == "Fighting"),
    cat_fighting = as.numeric(replace_na(cat_fighting, FALSE)),
    
    cat_science_fiction = if_any(starts_with("cate"), ~ .x == 
      "Science Fiction"),
    cat_science_fiction = as.numeric(replace_na(cat_science_fiction, FALSE)),
    
    cat_dice = if_any(starts_with("cate"), ~ .x == "Dice"),
    cat_dice = as.numeric(replace_na(cat_dice, FALSE)),
    
    cat_bluffing = if_any(starts_with("cate"), ~ .x == "Bluffing"),
    cat_bluffing = as.numeric(replace_na(cat_bluffing, FALSE)),
    
    cat_adventure = if_any(starts_with("cate"), ~ .x == "Adventure"),
    cat_adventure = as.numeric(replace_na(cat_adventure, FALSE)),
    
    cat_miniatures = if_any(starts_with("cate"), ~ .x == "Miniatures"),
    cat_miniatures = as.numeric(replace_na(cat_miniatures, FALSE)),
    
    cat_ww2 = if_any(starts_with("cate"), ~ .x == "World War II"),
    cat_ww2 = as.numeric(replace_na(cat_ww2, FALSE)),
    
    cat_medieval = if_any(starts_with("cate"), ~ .x == "Medieval"),
    cat_medieval = as.numeric(replace_na(cat_medieval, FALSE)),
    
    cat_explor = if_any(starts_with("cate"), ~ .x == "Exploration"),
    cat_explor = as.numeric(replace_na(cat_explor, FALSE)),
    
    cat_deduct = if_any(starts_with("cate"), ~ .x == "Deduction"),
    cat_deduct = as.numeric(replace_na(cat_deduct, FALSE)),
    
    cat_party = if_any(starts_with("cate"), ~ .x == "Party Game"),
    cat_party = as.numeric(replace_na(cat_party, FALSE)),
    
    cat_abstract = if_any(starts_with("cate"), ~ .x == "Abstract Strategy"),
    cat_abstract = as.numeric(replace_na(cat_abstract, FALSE)),
    
    cat_animals = if_any(starts_with("cate"), ~ .x == "Animals"),
    cat_animals = as.numeric(replace_na(cat_animals, FALSE))
  ) %>%
  # flags for mechanic
  step_mutate(
    mech_dice = as.numeric(str_detect(mechanic, "Dice Rolling")),
    mech_hand = as.numeric(str_detect(mechanic, "Hand Management" )),
    mech_powers = as.numeric(str_detect(mechanic, "Variable Player Powers")),
    mech_sets = as.numeric(str_detect(mechanic, "Set Collection")),
    mech_infl = as.numeric(str_detect(mechanic, 
      "Area Control / Area Influence")),
    mech_draft = as.numeric(str_detect(mechanic, "Card Drafting")),
    mech_modular = as.numeric(str_detect(mechanic, "Modular Board")),
    mech_tile = as.numeric(str_detect(mechanic, "Tile Placement")),
    mech_hex = as.numeric(str_detect(mechanic, "Hex-and-Counter")),
    mech_action = as.numeric(str_detect(mechanic, 
      "Action Point Allowance System")),
    mech_coop = as.numeric(str_detect(mechanic, "Co-operative Play")),
    mech_sas = as.numeric(str_detect(mechanic, 
      "Simultaneous Action Selection")),
    mech_auction = as.numeric(str_detect(mechanic, "Auction/Bidding")),
    mech_area = as.numeric(str_detect(mechanic, "Area Movement")),
    mech_worker = as.numeric(str_detect(mechanic, "Worker Placement")),
    mech_grid = as.numeric(str_detect(mechanic, "Grid Movement")),
    mech_simul = as.numeric(str_detect(mechanic, "Simulation")),
    mech_deck = as.numeric(str_detect(mechanic, "Deck / Pool Building")),
    mech_partner = as.numeric(str_detect(mechanic, "Partnerships")),
    mech_point = as.numeric(str_detect(mechanic, "Point to Point Movement")),
    mech_route = as.numeric(str_detect(mechanic, "Route/Network Building"))  
  ) %>%
  # flags for mechanic
  step_mutate(
    design_1 = as.numeric(str_detect(designer, "Reiner Knizia")),
    design_2 = as.numeric(str_detect(designer, "Martin Wallace")),
    design_3 = as.numeric(str_detect(designer, "Wolfgang Kramer")),
    design_4 = as.numeric(str_detect(designer, "Dean Essig")),
    design_5 = as.numeric(str_detect(designer, "Alan R. Moon")),
    design_6 = as.numeric(str_detect(designer, "Bruno Cathala")),
    design_7 = as.numeric(str_detect(designer, "Friedemann Friese")),
    design_8 = as.numeric(str_detect(designer, "Mike Elliott")),
    design_9 = as.numeric(str_detect(designer, "Klaus Teuber")),
    design_10 = as.numeric(str_detect(designer, "Richard H. Berg"))
  ) %>%
  # remove the name, designer, mechanic and category variables
  step_rm(names, mechanic, designer, category1:category12) %>%
  # log transformation
  step_log(num_votes, owned) %>%
  # outliers
  step_mutate(
    min_players = case_when(
      min_players == 0 ~ as.numeric(NA), 
      min_players > 20 ~ 20,
      TRUE ~ min_players
    ),
    max_players = case_when(
      max_players == 0 ~ as.numeric(NA), 
      max_players > 20 ~ 20,
      TRUE ~ max_players
    ),
    min_time = case_when(
      min_time == 0 ~ as.numeric(NA),
      min_time > 240 ~ 240,
      TRUE ~ min_time
    ),
    max_time = case_when(
      max_time == 0 ~ as.numeric(NA),
      max_time > 240 ~ 240,
      TRUE ~ max_time
    ),
    avg_time = case_when(
      avg_time == 0 ~ as.numeric(NA),
      avg_time > 240 ~ 240,
      TRUE ~ avg_time
    ),
    year = case_when(
        year == 0 ~ as.numeric(NA), 
        year < 1900 ~ 1900,
        TRUE ~ year
    )
  ) %>%
  # imputation
  step_impute_mean(min_players, max_players, year, min_time, max_time, avg_time)
```

# Check recipe outputs

```{r}
prepped <- prep(rec)
baked <- bake(prepped, new_data = NULL) 

skim(baked)  
```

# Specify model

```{r}
spec <- rand_forest(mtry = tune(), min_n = tune(), trees = 1500) %>%
  set_mode("regression") %>%
  set_engine("ranger")
```

# Specify workflow

```{r}
wkflow <- workflow() %>%
  add_recipe(rec) %>%
  add_model(spec)

wkflow  
```

# Tune

```{r}
set.seed(530)
train_folds <- vfold_cv(train)

set.seed(3621)
doParallel::registerDoParallel()
ranger_tune <-
  tune_grid(wkflow,
    resamples = train_folds,
    grid = 11
  )

show_best(ranger_tune, metric = "rmse") %>%
  kable( 
    escape = FALSE,
    padding = 5
  ) %>% 
  kable_styling(
    bootstrap_options = c("striped", "hover"), 
    fixed_thead = TRUE, 
    position = "left"
  ) %>%
  row_spec(0, color = "white", background = table_colour)

autoplot(ranger_tune) +
  theme_bw()
```

# Finalise model

```{r}
final_wkflow <- wkflow %>%
  finalize_workflow(select_best(ranger_tune, metric = "rmse"))

final_wkflow
```

# Fit model

```{r}
ranger_fit <- last_fit(final_wkflow, split)
```

# Check performance

```{r}
collect_metrics(ranger_fit) %>%
  kable( 
    escape = FALSE,
    padding = 5
  ) %>% 
  kable_styling(
    bootstrap_options = c("striped", "hover"), 
    fixed_thead = TRUE, 
    position = "left"
  ) %>%
  row_spec(0, color = "white", background = table_colour)
```

# Get predictions

```{r}
ranger_preds <- collect_predictions(ranger_fit)

ranger_preds %>%
  ggplot(aes(geek_rating, .pred)) +
  geom_abline(lty = 2, colour = pal_sliced[3]) +
  geom_point(alpha = 0.5) +
  coord_fixed() +
  labs(title = 'Predicted vs actual')
```

# Apply to new data

```{r}
have_scored <- cbind(to_score, predict(ranger_fit$.workflow[[1]], to_score)) %>%
  select(game_id, geek_rating = .pred)
```

# Output submission

```{r}
write_csv(have_scored, 'sliced/s01e01/submission.csv')
```

# Outcome

While this was a late submission, the RMSE for my submission was **0.17972**, which would put me about 10^th^ on the private leaderboard

# Possible improvements

- Try different model types
- Is there a better way to create flags for the cateogries and mechanics
